---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

# Please deposit as `assignment6` as a Readme.Rmd file.

This homework will generally follow: https://sites.usc.edu/translationalgenomics/learning-r/data-wrangling/

## Descriptive Statistics

### Getting Dataset 1

```{r}
clinical_info <- read.csv('trgn510/trgn599.clinical.tsv', header=TRUE, sep="\t")
rna <- read.csv('trgn510/trgn599.rna.tsv', header=TRUE, sep="\t")
```


## Mean, etc

We can caculate the mean 

```{r}
mean(clinical_info$height, na.rm = TRUE)
```

Standard Deviation

```{r}
sd(clinical_info$height, na.rm = TRUE)
```

Now, using aggregates:

```{r}
aggregate(clinical_info$cigarettes_per_day,by = list(race = clinical_info$race), na.rm=TRUE,mean)
```

###  Aggregate

Looking at two different variables using aggregate

```{r}
aggregate(clinical_info$cigarettes_per_day,by = list(race = clinical_info$race,ethnicity=clinical_info$ethnicity), na.rm=TRUE,mean)
```

### Another approach using %>%
```{r}
clinical_info %>%
  filter(Sample.Type == "Primary Tumor") %>%
  group_by(race) %>%
  summarise(mean(cigarettes_per_day,na.rm=TRUE))
```

Another verb is summarise

## summarise

```{r}
summarise(clinical_info$cigarettes_per_day)
```


## Group_by

group_by() is a very useful verb; as the name implies, it allows you to create groups and then, for example, compute descriptive statistics by groups. 

```{r}
c<-clinical_info %>% group_by(race)
```

While nothing appears to happen, look at the last line - the power comes when adding functions



### mutate

```{r}
clinical_info %>%
  transmute(years_to_death = exp(days_to_death /365))
```

          
          
## Downloading data and loading it

The commands below will download and create the files in our current path.

Make sure you have done `install.packages` of any missing packages, such as `install.packages('googledrive')`

```{r}
library(googledrive)
drive_deauth()
drive_download("https://drive.google.com/file/d/11ebdJSErJN58THWvlaiu8k7AKiMjHMFL/view?usp=sharing", type = "csv")
drive_download("https://drive.google.com/file/d/11cxySFxeF3SrJpRxkyP4YOQBn2rsIo-q/view?usp=sharing", type = "csv")

```

## Loading Data into a dataframe

```{r}
cancer_by_county <- read.csv('cancer_by_county.csv', header=TRUE, sep=",", strip.white=TRUE)
medicare_by_county <- read.csv('medicare_by_county.csv', header=TRUE, sep=",")
```

### Inspecting cancer_by_county
```{r}
head(cancer_by_county)
```
### Inspecting cancer_by_county
```{r}
head(medicare_by_county)
```


## Splitting

Split medicare by county "county.name" field into two new fields: "County" and "StateCode", seperating by "_" to give us a column 'StateCode' and 'County'

```{r}
library(tidyr)
library(stringr)
library(dplyr)
medicare_by_county['StateCode'] <- data.frame(do.call('rbind', strsplit(as.character(medicare_by_county$StateCode_County),'_')))[,1]
medicare_by_county['County'] <- data.frame(do.call('rbind', strsplit(as.character(medicare_by_county$StateCode_County),'_')))[,2]

```

Lets inspect
```{r}
head(medicare_by_county['County'])
```

Please join the data using DPLYR using an innerjoin
```{r}
library(dplyr)
cancer_medicare_by_county <- inner_join(medicare_by_county, cancer_by_county,"County")
```

Lets examine the resulting table:
```{r}
summary(cancer_medicare_by_county)
```



## Problems:

### Problem 1

Please save the joined_data to a file called "cancer_medicare_by_county" as a csv, without using rownames.

```{r}
# Place your code here.
write.csv(cancer_medicare_by_county, "CancermodifiedByCounty.csv",row.names = FALSE)
```

### Problem 2

What is the mean of Hospice.reimbursements.per.enrollee

```{r}
mean(cancer_medicare_by_county$Hospice.reimbursements.per.enrollee)
```

### Problem 3

Descriptive statistitics

aggregate(starwars$height,
          by = list(Species = starwars$species),
          mean)
          

```{r}
states<-aggregate(Hospice.reimbursements.per.enrollee$Hospice.reimbursements.per.enrollee,
          by = list(Species = Hospice.reimbursements.per.enrollee$StateCode),
          mean)
```


What is the mean of Medicare.enrollees?



